StreamReduce: Processing Data Streams\\and Avoiding Work

### 2012-05-07 paper/paper.tex
More edits

The remainder of this paper is organized as follows. Section~\ref{sec:relwork} provides
an overview of related work on distributed stream processing as well as dynamic and
approximate computations. Section~\ref{sec:design} details how I designed StreamReduce.
Section~\ref{sec:methodology} explains the methodology used to perform a case study
using StreamReduce. Section~\ref{sec:evaluation} evaluates the effectiveness of StreamReduce in the case study.
Section~\ref{sec:discussion} enumerates issues with StreamReduce's current implementation.

@@

@@

\section{Design}
\label{sec:design}
StreamReduce is a distributed system for processing streams that can choose to kill tasks or
limit the amount of work it performs per task. Its design is modular and there are five
different types of nodes in an StreamReduce cluster: spout, fetcher, worker, collector, master.
\subsection{Spout}
Spouts are network-connected machines that generate data for an StreamReduce cluster to process.
Examples include a machine that tails a log file or an API like the Twitter tweet
stream. Fetchers depend on spouts as a source of streaming data.
\subsection{Fetcher}
Fetchers are responsible for communicating with spouts and returning data for the
workers to process. Fetchers may run indefinitely or return a status code indicating
the fetching task is over.
\subsection{Worker}
Workers perform processing on data provided by fetchers. They are similar to mappers
in the MapReduce framework. Each worker receives a fraction of the total input data
and each worker's tasks are independent of other workers'.

Workers are responsible for fulfilling the main design goals of StreamReduce. Workers may choose
to kill a task preemptively or greedily if it does not complete before a timeout. Additionally,
workers may choose among a variety of routines for performing computation on a datum.
Routines are ranked on a [0,1] scale based on the estimated fraction of total work it
requires.

@@

Because one of StreamReduce's goals is to process (potentially infinite) streams, workers do not
save their progress to disk and are forgetful; once the master queries a worker for its
progress and the worker successfully returns it, the worker clears all of its state and
continues processing. As a consequence, workers are easy to add or remove from the worker
pool because they don't need to be initialized with long-term job state. Additionally, as
long as the master polls workers frequently enough, when a worker fails, only a small
fraction of the total work is lost. The master queries workers every 100ms by default.
\subsection{Collector}
Collector nodes aggregate the results of all workers. Each collector acts as a reducer would
in the MapReduce framework if MapReduce used a single reducer for all of the mappers.
Collectors differ from MapReduce reducers in that they receive their data piecewise, as it
is produced from workers as opposed to after all mappers have completed. This limits the
types of stateless aggregation collectors can do to online operations. (Of course, a collector
can choose to retain copies of individual worker results and re-reduce them each time more
results are receieved, but this is not in the spirit of the framework.)

@@

StreamReduce achieves this configuration with a Jobfile. Jobfiles are instances of a special class
provided by the StreamReduce framework. This class exposes several methods that can be invoked by a
given node to provide configuration parameters or execute code on its behalf. Most of the
routines that advance the progress of an StreamReduce job are wrappers around these Jobfile methods.

@@

The use of eval to instantiate Jobfile objects and the implementation language's (Ruby)
ability to monkeypatch classes means that jobs can be modified on the fly just by sending
an updated Jobfile to job participants. For example, a job can be made lower fidelity in
the face of an impending time constraint, higher fidelity if more workers are added to
the cluster, or gracefully killing fetchers which allows the job to end after all of the
various job queues are empty.
\subsection{Fault Tolerance}
StreamReduce can sustain the failure of any node other than the master. Because the master is responsible
for shuffling data between fetchers, workers, and collectors, a job cannot make progress without
the master.

Fetchers can fail. The only consequence of fetcher failure is that some of the input stream is
lost forever. Since streams contain much data and are possibly stored permanently elsewhere on disk
StreamReduce does not attempt to recover this data.

@@

These behaviors appeal to the law of large numbers. It assumes that trends will still be
visible in spite of looking at only a fraction of the data. This means that the computation
routines and the weights assigned to them must be adapted to fit the distribution of the data
before the job starts (or updated in situ via a new Jobfile).

@@

\section{Evaluation}
\label{sec:evaluation}
To evaluate StreamReduce, I explore the precision and recall of the word count test job as well as the
speedup factor associated with varying the kill rate and target work level.

@@

\subsubsection{Precision}
Task killing should least impact the head of the word distribution.
Figure~\ref{fig:precision} shows that even with a kill rate of~0.8 and a work level of~0.4,
meaning that the cluster only looks at 8\% of the words in the corpus, the results
are at least~80\% accurate. Increasing the work level, which increases the fidelity of
results emitted by workers, also increases the fidelity of the top~100 results.

\begin{figure}
\includegraphics[width=\linewidth]{top-100-precision.png}
\caption{Fraction of the top~100 results returned in the top-100 by the word count job.}
\label{fig:precision}
\end{figure}

\subsubsection{Word Count Job Accuracy}
Because I am running a top-k job, I compute the $F_{0.5}$-score of the StreamReduce results as
opposed to the $F1$-score because I care more about precision of the top k than recall
in the tail. Figure~\ref{fig:fscore} shows that the $F_{0.5}$-score, or the weighted
accuracy of the top-100 classifier, is always higher than the fraction of the input
examined (kill rate * work level). Setting the work level to~0.4 nets a greater increase
in job accuracy versus fraction of words examined, with accuracy approximately~1.75 times
more than the fraction of words examined.

@@

Table~\ref{table:runtime} shows the effect of varying the kill rate and work level. The row
with kill~rate~$=0.0$ and work~level~$=1.0$ represents the perfect, MapReduce-style, execution
of the word count job. The only consistent reduction in execution time is a result of
increasing the kill rate. Increasing the kill rate exhibits a linear reduction in execution
time as the fraction of processed data linearly decreases.

\begin{table}
\begin{tabularx}{\linewidth}{|c|c|X|}
\hline
Kill rate & Work level & Average execution time [minutes] \\ \hline
0.0 & 1.0 & 11:35.5 \\ \hline
0.0 & 0.4 & 10:33.8 \\ \hline
0.0 & 0.8 & 10:30.8 \\ \hline
0.2 & 0.4 & 08:25.8 \\ \hline
0.2 & 0.8 & 08:37.8 \\ \hline
0.4 & 0.4 & 06:54.0 \\ \hline
0.4 & 0.8 & 06:28.2 \\ \hline
0.6 & 0.4 & 04:14.7 \\ \hline
0.6 & 0.8 & 04:13.7 \\ \hline
0.8 & 0.4 & 02:33.3 \\ \hline
0.8 & 0.8 & 02:15.0 \\ \hline
\end{tabularx}
\caption{Impact of kill rate and work level on job execution time.}
\label{table:runtime}
\end{table}

>>>>

The remainder of this paper is organized as follows. Section~\ref{sec:relwork} provides
an overview of related work on distributed stream processing as well as dynamic and
approximate computations. Section~\ref{sec:design} details the design of StreamReduce.
Section~\ref{sec:methodology} explains the methodology used to perform a case study
using StreamReduce. Section~\ref{sec:evaluation} evaluates the effectiveness of StreamReduce in the case study.
Section~\ref{sec:discussion} enumerates issues with StreamReduce's current implementation.

@@

Aurora also introduces the operators \texttt{filter}, \texttt{map}, and \texttt{union} for
manipulating streams. StreamReduce implements these operators at the worker and fetcher nodes.

@@

\section{Design}
\label{sec:design}
StreamReduce is a distributed system for processing streams that can choose to kill tasks or
limit the amount of work it performs per task. Its design is modular and there are five
different types of nodes in a StreamReduce cluster: spout, fetcher, worker, collector, master.
\subsection{Spout}
Spouts are network-connected machines that generate data for a StreamReduce cluster to process.
Examples include a machine that tails a log file or an API like the Twitter tweet
stream. Fetchers depend on spouts as a source of streaming data.
\subsection{Fetcher}
Fetchers are responsible for communicating with spouts and returning data for the
workers to process. This level of indirection allows for a variety of different
configurations: A spout can emit data at such a high rate that using multiple fetchers
returns more data to the cluster, multiple fetchers can each talk to a different spout,
for example tailing logs from multiple machines, or fetchers can be set up redundantly
to ensure that every datum emitted by a spout is processed.

Fetchers may run indefinitely or return a status code indicating
the fetching task is over.
\subsection{Worker}
Workers perform processing on data provided by fetchers. They are similar to mappers
in the MapReduce framework. Each worker receives a fraction of the total input data
and each worker's tasks are independent of other workers'.

Workers are responsible for fulfilling the main design goals of StreamReduce. Workers may choose
to kill a task preemptively or, if it does not complete before a timeout, greedily. Additionally,
workers may choose among a variety of routines for performing computation on a datum.
Routines are ranked on a [0,1] scale based on the estimated fraction of total work it
requires. This rank is a proxy for the accuracy of a routine.

@@

Because one of StreamReduce's goals is to process (potentially infinite) streams, workers do not
save their progress to disk and are forgetful; once the master queries a worker for its
progress and the worker successfully returns it, the worker clears all of its state and
continues processing. As a consequence, workers are easy to add or remove from the worker
pool because they don't need to be initialized with long-term job state. Additionally, as
long as the master polls workers frequently enough, only a small
fraction of the total work is lost upon worker failure.
The master queries workers every 100ms by default.
\subsection{Collector}
Collector nodes aggregate the results of all workers. Each collector acts as a reducer would
in the MapReduce framework if MapReduce used a single reducer for all of the mappers.
Collectors differ from MapReduce reducers in that they receive their data piecewise, as it
is produced from workers, as opposed to after all mappers have completed. This limits the
types of stateless aggregation collectors can do to online operations. (Of course, a collector
can choose to retain copies of individual worker results and re-reduce them each time more
results are receieved, but this is not in the spirit of the framework.)

@@

StreamReduce achieves this configuration with a Jobfile. Jobfiles are instances of a special class
provided by the StreamReduce framework. This class exposes several methods that can be invoked by a
given node to provide configuration parameters or execute code on its behalf. Most of the
routines that advance the progress of a StreamReduce job are wrappers around these Jobfile methods.

@@

The use of eval to instantiate Jobfile objects and the implementation language's (ruby)
ability to monkeypatch classes means that jobs can be modified on the fly just by sending
an updated Jobfile to job participants. For example, a job can be made lower fidelity in
the face of an impending time constraint, be made higher fidelity if more workers are added to
the cluster, or gracefully kill all fetchers which allows the job to end after all of the
various job queues are empty.
\subsection{Fault Tolerance}
StreamReduce can sustain the failure of any node other than the master. Because the master is responsible
for shuffling data between fetchers, workers, and collectors, a job cannot make progress without
the master.

Fetchers can fail. The only consequence of fetcher failure is that some of the input stream is
lost forever. Because StreamReduce uses the Data Stream Model, it does not attempt to recover this data.

@@

These behaviors appeal to the law of large numbers. It assumes that trends will still be
visible in spite of looking at only a fraction of the data. This means that the computation
routines and the weights assigned to them must be adapted to fit the distribution of the data
before the job starts (or updated \emph{in situ} via a new Jobfile).

@@

\section{Evaluation}
\label{sec:evaluation}
To evaluate StreamReduce, I explore the precision and recall of the word count test job as well as the
speedup factor associated with varying the task kill rate and target work level.

@@

\subsubsection{Precision}
Task killing should least impact the head of the word distribution.
Figure~\ref{fig:precision} shows that even with a task kill rate of~0.8 and a work level of~0.4,
meaning that the cluster only looks at 8\% of the words in the corpus, the results
are at least~80\% accurate. Increasing the work level, which increases the fidelity of
results emitted by workers, also increases the fidelity of the top~100 results.

\begin{figure}
\includegraphics[width=\linewidth]{top-100-precision.png}
\caption{Fraction of the top~100 results returned in the top~100 by the word count job.}
\label{fig:precision}
\end{figure}

\subsubsection{Word Count Job Accuracy}
Because I am running a top-k job, I compute the $F_{0.5}$-score of the StreamReduce results as
opposed to the $F1$-score because I care more about precision of the top k than recall
in the tail. Figure~\ref{fig:fscore} shows that the $F_{0.5}$-score, or the weighted
accuracy of the top-100 classifier, is always higher than the fraction of the input
examined (task kill rate * work level). Setting the work level to~0.4 nets a greater increase
in job accuracy versus fraction of words examined, with accuracy approximately~1.75 times
higher than the fraction of words examined.

@@

Table~\ref{table:runtime} shows the effect of varying the task kill rate and work level. The row
with kill~rate~$=0.0$ and work~level~$=1.0$ represents the perfect, MapReduce-style, execution
of the word count job. The only consistent reduction in execution time is a result of
increasing the task kill rate. Increasing the task kill rate exhibits a linear reduction in execution
time as the fraction of processed data linearly decreases.

\begin{table}
\begin{tabularx}{\linewidth}{|c|c|X|}
\hline
Task kill rate & Work level & Average execution time [minutes] \\ \hline
0.0 & 1.0 & 11:35.5 \\ \hline
0.0 & 0.4 & 10:33.8 \\ \hline
0.0 & 0.8 & 10:30.8 \\ \hline
0.2 & 0.4 & 08:25.8 \\ \hline
0.2 & 0.8 & 08:37.8 \\ \hline
0.4 & 0.4 & 06:54.0 \\ \hline
0.4 & 0.8 & 06:28.2 \\ \hline
0.6 & 0.4 & 04:14.7 \\ \hline
0.6 & 0.8 & 04:13.7 \\ \hline
0.8 & 0.4 & 02:33.3 \\ \hline
0.8 & 0.8 & 02:15.0 \\ \hline
\end{tabularx}
\caption{Impact of task kill rate and work level on job execution time.}
\label{table:runtime}
\end{table}
